Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-08-17 paper_comparison/paper_comparison/paper_comparison.tex
intro restructured

% see https://stackoverflow.com/a/47122900

% Pandoc citation processing

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\title{Effectiveness of visual test in linear regression diagnositics}

@@

\begin{keywords}
visual inference; regression diagnostics;
\end{keywords}

@@

Regression diagnostics is an essential step in regression analysis which
is a field of study with at least a hundred years of history. The
diagnostic procedure conventionally involves evaluating the fitness of
the proposed model, detecting the presence of influential observations
and outliers, checking the validity of model assumptions and many more.
In terms of diagnostic techniques, data plots, hypothesis testing, and
summary statistics are vital tools for a systematic and detailed
examination of the regression model \citep{mansfield1987diagnostic}.

Many of those regression diagnostic methods and procedures are mature
and well-established in books first published in the twentieth century,
such as \citet{draper1998applied}, \citet{montgomery1982introduction},
\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
\citet{cook1982residuals}. Regardless of the level of difficulty, one
will find the importance and usefulness of diagnostic plots being
emphasized in those books repeatedly. Checking diagnostic plots is also
the recommended starting point for validating model assumptions such as
normality, homoscedasticity and linearity
\citep{anscombe_examination_1963}.

@@

\begin{figure}
\centering
\includegraphics[width=4.6875in,height=3.55208in]{figures/rsta2009012001.jpg}
\caption{Parallelism between multiple quantitative testing and visual
discovery \citep{buja_statistical_2009}. Visible features in a plot are
viewed as a collection of test statistics
\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\), and any visual
discoveries that are inconsistent with the null hypothesis are treated
as evidence against the null. For regression diagnostics, the null
hypothesis would be the assumed model, and visual discoveries could be
any visual features in favour of any alternatives.
\label{fig:parallelism}}
\end{figure}

>>>>

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\title{Why aren't significance tests commonly used for linear regression
diagnostics?}

@@

\begin{keywords}
data visualization; visual inference; hypothesis testing; residual
plots;
\end{keywords}

@@

\begin{quote}
\emph{``Since all models are wrong the scientist must be alert to what
is importantly wrong.''} \citep{REF}
\end{quote}

Diagnosing a model is the key to determining whether there is anything
importantly wrong. For linear regression analysis, it is typical to
interrogate the residuals. Residuals summarise what is not captured by
the model, and thus provide the capacity to identify what might be
wrong. There are many ways that residuals are assessed.

Residuals might be plotted, as a histogram or normal probability plot to
examine the distribution. Is it symmetric, and unimodal? Then all is
good. If the distribution is skewed or bimodal or multimodal, there are
outliers, there is cause for concern. It if also possible to conduct a
goodness of fit test, such as a Shapiro-Wilk Normality test \citep{REF}.

It is also recommended to plot residuals against predicted values, and
also each of the explanatory variables. If there is no obvious pattern,
then all is good. It is a very difficult task for a human judge, though
to make a decision that there's nothing there. It is especially common,
particularly among new data analysts to report patterns when an
experienced data analyst might quickly conclude that there are none.
Generally, one looks for departures from nothingness like nonlinear
dependency or heteroskedasticity. It is also possible to conduct a
hypothesis test for non-linear dependence \citep{RESET}, and also use a
Breusch-Pagan test \citep{REF} for heteroskedasticity.

There is an abundance of literature describing appropriate diagnostic
methods for linear regression: Cook and Weisberg, \ldots{} \citep{REF}.
The diligent reader of these sage writings will also notice sentences
that express sentiments like \emph{based on their experience,
statistical tests are not widely used in regression diagnostics. The
same or even larger amount of information can be provided by diagnostic
plots than the corresponding tests in most empirical studies.} There is
a common guidance by experts that plots are the best for diagnosing
model fits.

This is curious, and investigating why this might be common advice is
the subject of this paper. The paper is structured as follows. The next
background section describes the commonly used diagnostic plots, the
types of departures that one expects to detect, and describes a formal
process for reading residual plots, called visual inference, that can
avoid the concerns about subjectiveness of human readers. Section XXX
describes the experimental setup to enable a comparison between
decisopns made by formal hypothesis testing, and how humans would read
diagnostic plots. The results are reported in Section XXX. We finish
with a discussion on future work, in particular how the responsibility
for residual plot reading might be passed on to computer vision.

\hypertarget{background}{%
\section{Background}\label{background}}

@@
